---
title: "Wheat Harvested Surface in Spain 2004"
author: "Stephen Blatti"
date: "October 7, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

6. The wheat table shown on p26 gives the wheat harvested surfaces in 2004 by autonomous communities in Spain measured in hectares. 
```{r}
library(PASWR2)
```
a) Create the var's community and wheat.surface from the Wheat Table. store both var's in a data.frame named wheat.spain.
```{r}
community <- c("Galicia", "Asturias", "Cantabria", "Pais Vasco", "Navarra", "La Rioja", "Aragon", "Cataluna", "Islas Baleares", "Castilla y Leon", "Madrid", "Castilla-La Mancha", "C. Valenciana", "Region de Murcia", "Extremadura", "Andalucia", "Islas Canarias")
wheat.surface <- c(18817, 65, 440, 25143, 66326, 34214, 311479, 74206, 7203, 619858, 13118, 263424, 6111, 9500, 143250, 558292, 100)
wheat.spain <- data.frame(community, wheat.surface)
head(wheat.spain)
```
b) Find the maximum, the minimum, and the range for the variable wheat.surface.
```{r}
max(wheat.spain$wheat.surface)
min(wheat.spain$wheat.surface)
diff(range(wheat.spain$wheat.surface))
```
c) Which community has the largest harvested wheat surface?
```{r}
wheat.spain[wheat.spain$wheat.surface == max(wheat.spain$wheat.surface),]
```
d) Sort the autonomous communities in ascending order.
```{r}
asc_order <- wheat.spain[order(wheat.spain$wheat.surface, decreasing = FALSE), ]
head(asc_order)
```
e) Sort the autonomous communities in decending order.
```{r}
dec_order <- wheat.spain[order(wheat.spain$wheat.surface, decreasing = TRUE), ]
head(dec_order)
```
f) Create a new file called wheat.c where Asturias has been removed.
```{r}
wheat.c <- wheat.spain[wheat.spain$community != "Asturias", ]
head(wheat.c)
```
g) Add Asturias back to the file wheat.c
```{r}
wheat.c <- rbind(wheat.c, wheat.spain[wheat.spain$community == "Asturias", ])
wheat.c
```
h) Create in wheat.c a new variable called acre indicating the harvested surface in acres (1 acre = 0.40468564224 hectares).
```{r}
wheat.c <- within(data = wheat.c, { 
  acre <- wheat.surface/0.40468564224
  })
head(wheat.c)
```
i) What is the total harvested surface in hectares and in acres in Spain in 2004?
```{r}
sum(wheat.c$wheat.surface)
sum(wheat.c$acre)
```
j) Define in wheat.c the row.names() using the names of the communities. Remove the community variable from wheat.c
```{r}
#rm(community, wheat.surface)
no_community <- wheat.c[, -1]
row.names(no_community) <- wheat.c[, 1]
wheat.c <- no_community
head(wheat.c)
```
k) What percent of the commumities have a harvested wheat surface greater than the mean wheat surface area? 
```{r}
PCA <- mean(wheat.c$wheat.surface > mean(wheat.c$wheat.surface))*100
PCA
```
l) Sort wheat.c by autonomous communities' name (row.names()).
```{r}
sorted_AO <- wheat.c[order(row.names(wheat.c)), ]
head(sorted_AO)
```
m) Determine the communities with less than 40,000 acres of harvested surface and find thier total harvested surface in hectares and acres.
```{r}
less_than_40k <- wheat.c[wheat.c$acre < 40000, ]
less_than_40k
apply(less_than_40k, 2, sum)
```
n) Create a new file called wheat.sum where the autonomous communities that have less than 40,000 acres of harvested surface are consolidated into a single category named "less than 40,000" with the results from (m).
```{r}
lt_40 <- apply(less_than_40k, 2, sum)
gt_40 <- wheat.c[wheat.c$acre >= 40000, ]
wheat.sum <- rbind(gt_40, lt_40)
row.names(wheat.sum)[11] <- c("less than 40,000")
wheat.sum
```
o) Use the function dump() on wheat.c storing the results in a new file named wheat.txt. Remove wheat.c from your path and check that you can recover it from wheat.txt.
```{r}
dump("wheat.c", file = "wheat.txt")
rm("wheat.c")
source("wheat.txt")
head(wheat.c)

```
p) Create a text file called wheat.dat from the wheat.sum file using the command write.table(). Explain the differences between wheat.txt and wheat.dat.
```{r}
write.table(x = wheat.sum, file = "wheat.dat")
# The values from wheat.txt are collapsed down in wheat.dat
```
q) Use the comand read.table() to read the file wheat.dat
```{r}
head(read.table(file = "wheat.dat"))
```













