---
title: "Problem 2.5"
author: "Stephen Blatti"
date: "October 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
2.5 The data frame vit2005 in the PASWR package contains descriptive info and the appraised total price (in euros) for appartments in Victoria, Spain.
```{r}
library(PASWR2)
```
a) Create a freq table, piechart, and a barplot showing the num of apartments grouped by the var out. Which method conveys the info best?
```{r}
VIT2005$out <- factor(VIT2005$out, levels = c("E25", "E50", "E75", "E100"))
plot1 <- ggplot(data = VIT2005, aes(x = factor(1), fill = out))+ geom_bar(width = 1) + coord_polar(theta = "y") + theme_bw() + labs(x = "", y = "")
plot2 <- ggplot(data = VIT2005, aes(x = out)) + geom_bar(fill = "lightblue") + theme_bw()
plot1
plot2
# The barplot is best
```
b) Charaterize the distribution of the var totalprice. 
```{r}
ggplot(data = VIT2005, aes(x = totalprice)) + geom_histogram(fill = "lightblue") + theme_bw()
max(VIT2005$totalprice) #one outlier at 560000.
median(VIT2005$totalprice)
IQR(VIT2005$totalprice)
# The distribution of totalprice is skewed to the right. The median totalprice is 269750 and the IQR for totalprice is 100125.
```
c) Charaterize the relationship between totalprice and area.
```{r}
ggplot(data = VIT2005, aes(x = area, y = totalprice)) + geom_point() + theme_bw()
# The relationship between totalprice and area is of positive linear.
```
d) Create a Trellis plot of totalprice vs area conditioning on toilets. Between what 2 values of area do apartments have both one and two bathrooms?
```{r}
xyplot(totalprice ~ area | toilets, data = VIT2005, layout = c(1, 2), as.table = TRUE)
ggplot(data = VIT2005, aes(x = area, y = totalprice, color = as.factor(toilets))) + geom_point() + facet_grid(toilets ~ .) + theme_bw() +  guides(color = guide_legend("Number of\nToilets"))
# Apartments with one bathroom: approx between 50 and 100 m^2
# Apartments with two bathrooms: approx between 80 and 120 m2. 
# The intersection of apartments with one and two bathrooms is approximately (80 to 100)m^2.
```
e) Use the area values from part (d) to create a subset of apartments that have both 1 and 2 bathrooms. By how much does an additional bathroom increase the appraised value of the apartment?  Would you be willing to pay for an additioal bathroom if you lived in Spain?
```{r}
both_bathrooms <- subset(VIT2005, subset = area >= 80 & area <= 100)
result <- tapply(both_bathrooms$totalprice, both_bathrooms$toilets, median)
result
diff(result)
# The increase in totalprice for a second bathroom for apartments between
# 80 and 100 m^2 is 36000 euros - probably not willing to pay the addiional costs.
```






